A Comparsion of Traffic Engineering Methods for Application Perceived Metrics

### 2015-07-10 background-new.tex
Background: updated motivation, and Chapter intro subsection planning for Beyond MLU & NCDN

\chapter{Traffic engineering in content-dominated networks}
\label{ch:te-background}

This chapter serves two main purposes. First, is to provide the reader with necessary background on traffic engineering in ISPs, content delivery and prior work on the interaction between traffic engineering and content delivery. Second, is to explain the motivation for our work on traffic engineering in content-dominated networks in light of the existing work in this area.

To that end, this section has the following key objectives:
\begin{itemize}
	\item
	explain the common traffic engineering approaches and cost objectives in the context of an Internet service provider
	\item
	explain the content delivery techniques used by Content Deilvery Networks (CDNs) towards improving user-perceived performance.
	\item
	review prior literature on the interaction between traffic engineering and content delivery, in particular the interaction between overlay routing and network routing, and request redirection and network routing.
	\item
	make a case for studying the interaction between content placement and network routing by illustrating this interaction using examples. 
\end{itemize}

@@

\section{Traffic engineering}

@@

Traffic engineering schemes are commonly evaluated based on cost functions that are a function of utilization of network links. Link utilization is a relevant metric because a low utilization of all links implies that the network is free from congestion hotspots and has spare capacity to tolerate an increase in demand. A well-known cost function is maximum link utilization or MLU. We give a simple example to show how traffic engineering can reduce link utilization.

If the traffic matrix is known accurately, the optimal solution to the traffic engineering problem can formulated as a multi-commodity flow optimization problem. The routing thus computed is reffered to as \emph{Optimal TE} in literature. But, it is impossible to have accurately knowledge of future traffic matrices. Therefore, practical traffic engineering schemes either take a demand-aware or a demand-oblivious approach.  A demand-aware TE periodically updates routing using historically observed traffic matrices. A demand-oblivious TE requires no explicit measurement of traffic matices, but instead configures routing statically. A demand-oblivious TE is simpler to implement because it requires neither measuremnt nor periodic updating of network routing. But, demand-aware schemes have been shown to outperform demand-oblivious TE.

In practice, offline TE based on Open Shortest Path First (OSPF) and Multiprotocol Label Switching (MPLS) are commonly used  \cite{COPE,MultiTM,fortz2000internet,MPLS2}. Routes computed by OSPF traffic engineering must follow shortest-weight paths, therefore OSPF TE provides limited functionality to split traffic among multiple paths. MPLS TE overcomes this limitation by enabling traffic between two nodes to be split  in arbitrary ratios among multiple paths. Therefore, MPLS TE gives better results than OSPF TE as exemplified by the above example \cite{COPE,MultiTM}.

@@

\section{Content delivery}

@@

Strong consistency is a separate topic and is studied in the content of design of geo-distrbuted kv store in auspice.\tbd{fix this}

@@

Drawing parallels between our classification of traffic engineering schemes into demand-oblivious and demand-aware, we consider common content delivery techniques discussed above to be demand-oblivious since they do not require explicit measurement of content demand. In contrast, a demand-aware approach to content placement and redirection has been studied in research also, e.g., Applegate \emph{et al.} use a demand-aware approach to determine placement and redirection for Video-on-Demand content in the network.

\section{Interaction between traffic engineering and content delivery}

Content delivery routes traffic at the application layer using an overlay network of servers, and traffic enginering configures network routing of the underlying physical paths. Studying this interaction between decisions at the overlay and the underlying network has a topic of long standing interest in computer science. Several related questions have been put forth. Does the interaction between network routing and content delivery yield sub-optimal results? How can we design cooperative mechanisms to leverage these interactions? Do cooperative mechanisms yield benefit in an Internet-like environment? We find that nearly all previous research on this topic has focused either on the interaction of overlay routing with the network routing \cite{Roughgarden,selfishQiu}, or that of server selection strategies with the underlying routing  \cite{Jiang2009,JohariGameTheory, CATE, P4P} as discussed below.

@@

\section{Illustrating the interaction between content placement and traffic engineering}

\begin{figure}[tbh]
	\begin{center}
\label{fig:3node-bg}\includegraphics[scale=0.5]{final_images/Diagram3node.png}

	\caption{Lasso network}
	\end{center} 
	\end{figure}

%\textbf{explain conditions}
\textbf{Example 1:} In Figure~\ref{fig:3node-bg}, all links are assumed to have a capacity of 100 units and a constant delay. The top link A has a very small delay compared to the other two links that both have equal delay. Node 1 has 100 Mbps of demand that it can obtain from 2 as well as 3. In addition, there is 20 Mbps of demand at node 1 which it can obtain only from 2.  We assume that the aggregate demand at a node consists of a large number of user-initiated connections. When content can be downloaded from multiple locations, users initiate parallel TCP connections and the throughputs along paths in a parallel TCP connection are inversely proportional to the path delays. The TE scheme is assumed to be OSPF-based, i.e., shortest-path routing using configured link weights and traffic split equally among multiple paths with equal weights.

@@

\textbf{Example 2:}To appreciate how placement can shape traffic, consider the simple example in Figure~\ref{fig:NetworkExample}. Node $C$ has an object in its cache that is requested by end-users at nodes $A$ and $D$. Suppose that one unit of traffic needs to be routed from $C$ to $A$ and $0.5$ units  from $C$ to $D$ to satisfy the demand for that object. The routing that achieves the minimum MLU of $0.5$ to serve the demanded object is shown in the figure. Note that the routing that achieves the MLU of 0.5 is not possible with a simple, \unplanned\ protocol like InverseCap as that would route all the traffic demand from $C$ to $A$ via $B$, resulting in an MLU of 1. Thus, a (\planned) traffic engineering scheme is necessary to achieve an MLU of 0.5.

	\begin{figure}[h]
		\centering
		\includegraphics[width=2in]{ncdnpaper/ncdn-example}
		\caption{A Simple \ncp\ Example}
		\vspace{-.3in}
		\label{fig:NetworkExample}
	\end{figure}

@@

@@

>>>>

\chapter{Traffic engineering in content-dominated networks: background and motivation}
\label{ch:te-background}

This chapter serves two main purposes. First is to provide the reader with necessary background on ISP traffic engineering, content delivery and the interaction between them (Section \ref{sec:bg-bg}). Second is to explain the motivation for our work on traffic engineering in content-dominated networks in light of the existing work in this area (Section \ref{sec:bg-motivation}).

In comparison, following are the main new research directions that we pursue in this thesis. We explain each of them further in the chapter.

\begin{itemize}
	\item
	How do ISP traffic engineering schemes compare in terms of application-performance metrics such as TCP download time? Further, how do they compare while accounting fo r the effect of  application adaptation? 
	\item
	How should network CDNs--an ISP that operates a CDN on its infrastructure to deliver content to users on its network--perform content delivery given that they control both the content delivery and the underlying network routing?
	\item
	What role does content placement influence the interaction between network and content delivery, and what its its effect on user-perceived performance and netowrk cost?
\end{itemize}

@@

\section{Background}
\label{sec:bg-bg}
Our review of prior work provides following main findings:

\begin{itemize}
	\item
	ISP traffic engineering is studied mostly as an optimization problem for minimizing link-utilization based cost functions (Section \ref{sec:ch2-te}).
	\item
	Content delivery networks (CDNs) commonly use \emph{demand-oblivious} techniques for content placement and request redirection towards improving user-perceived performance across the Internet (Section \ref{sec:ch2-cdn}).
	\item
	Interaction between traffic engineering and content delivery has been shown to affects user-perceived performance as well as network cost. But, prior work has focused mostly on the interaction of traffic engineering with either overlay routing or with request redirection (Section \ref{sec:ch2-te-cdn}).
\end{itemize}

\subsection{Traffic engineering}
\label{sec:ch2-te}

@@

The evaluation metric for traffic engineering is a cost function that is dependent on the utilization of network links. Link utilization is a relevant metric because a low utilization of all links implies that the network is free from congestion hotspots and has spare capacity to tolerate an increase in demand. A well-known cost function is maximum link utilization or MLU. We give a simple example to show how traffic engineering can reduce link utilization.

We classify traffic engineering approaches into three groups:
\begin{description}
	\item[Optimal] If the traffic matrix is known accurately, the optimal solution to the traffic engineering problem can formulated as a multi-commodity flow optimization problem. The routing thus computed is reffered to as \textbf{Optimal} traffic engineering in literature.
	\item[Demand-aware] As it is impossible to have accurate knowledge of future traffic matrices, a demand-aware  traffic engineering  periodically updates routing using historically observed traffic matrices.
	\item[Demand-oblivious] A demand-oblivious  traffic engineering  requires no explicit measurement of traffic matices, but instead configures routing statically. A demand-oblivious  traffic engineering  is simpler to implement because it requires neither measuremnt nor periodic updating of network routing.

\end{description}

%As it is impossible to have accurately knowledge of future traffic matrices. Therefore, practical traffic engineering schemes either take a demand-aware or a demand-oblivious approach.  A demand-aware  traffic engineering  periodically updates routing using historically observed traffic matrices. A demand-oblivious  traffic engineering  requires no explicit measurement of traffic matices, but instead configures routing statically. A demand-oblivious  traffic engineering  is simpler to implement because it requires neither measuremnt nor periodic updating of network routing. But, demand-aware schemes have been shown to outperform demand-oblivious  traffic engineering.

In practice, demand-aware traffic engineering based on Open Shortest Path First (OSPF) and Multiprotocol Label Switching (MPLS) are commonly used  \cite{COPE,MultiTM,fortz2000internet,MPLS2}. Routes computed by OSPF traffic engineering must follow shortest-weight paths, therefore OSPF TE provides limited functionality to split traffic among multiple paths. MPLS TE overcomes this limitation by enabling traffic between two nodes to be split  in arbitrary ratios among multiple paths. Therefore, MPLS TE gives better results than OSPF TE as exemplified by the above example \cite{COPE,MultiTM}.

@@

\subsection{Content delivery}
\label{sec:ch2-cdn}

@@

%Strong consistency is a separate topic and is studied in the content of design of geo-distrbuted kv store in auspice.\tbd{fix this}

@@

Similar to our classification of traffic engineering schemes, we consider common content delivery techniques discussed above to be demand-oblivious since they do not require explicit measurement of content demand. In contrast, a demand-aware approach to content placement and redirection has also been studied, e.g., Applegate \emph{et al.} use a demand-aware approach to determine placement and redirection for Video-on-Demand content in the network.

\subsection{Interaction between traffic engineering and content delivery}
\label{sec:ch2-te-cdn}

Content delivery routes traffic at the application layer using an overlay network of servers, and traffic enginering configures network routing of the underlying physical paths. Studying this interaction between decisions at the overlay and the underlying network has been a topic of long standing interest in computer science. Several related questions have been put forth. Does the interaction between network routing and content delivery yield sub-optimal results? How can we design cooperative mechanisms to leverage these interactions? Do cooperative mechanisms yield benefit in an Internet-like environment? We find that nearly all previous research on this topic has focused either on the interaction of overlay routing with the network routing \cite{Roughgarden,selfishQiu}, or that of server selection strategies with the underlying routing  \cite{Jiang2009,JohariGameTheory, CATE, P4P} as discussed below.

@@

\section{Motivation}
\label{sec:bg-motivation}

The following key points summarize our motivation for studying traffic engineering in content-dominated netowrks.
\begin{itemize}
	\item
	Are link-utilization based metrics may good predictors of application performance that depends on other factors such as propagation delay, access and backbone link capacity, etc? Further, are these metrics are also poor predictors of capacity in scenarios when applications can download content from multiple locations?
	\item
	Since content placement can be a more powerful factor in reducing network cost, can content placement flexibility obviate the need for sophisticated traffic engineering?
	\item
	Network CDNs can control placement, redirection and routing on their network. Should they make these decisions independently or jointly? Further, should they use a demand-aware or a demand-oblivious approach to make these decisions?
\end{itemize}
%First, link-utilization based metrics may not be good predictors of application performance that depends on other factors such as queuing delay, access and backbone link capacity, etc. 
%Second, most work considers traffic engineering in isolation ignoring its interaction with content delivery, and even work 
%that accounts for this interaction has not focused on the role of placement strategies.
%Third, the problem of managing a network CDN in which a single entity handles both content delivery and traffic engineering has not been considered before.

\subsection{Do link-utilization based metrics reflect application performance and network capacity?}

Link-utilization based metrics may not be a good predictor of application performance.
Link-utilization primarily affects network quueing delay and loss. 
Application performance depends on other factors such as propagtion delay, access and backbone link capacity as well.
Queuing-theoretic models show that a high link utilization adversely affect queuing delay only at high utilization levels. For low to moderate utilization levels that are common in today's network, a reduction in link utilization may not yield a commensurate benefit in application peformance.

\begin{figure}[tbh]
	\centering
	\label{fig:3node-bg}\includegraphics[scale=0.5]{final_images/Diagram3node.png}
	\caption{Lasso network}
\end{figure}

Link-utilization based metrics may not be a good predictor of network capacity in scenarios where applications have the ability to download content from mulitple locations. In Figure~\ref{fig:3node-bg}, all links are assumed to have a capacity of 100 units and a constant delay. The top link A has a very small delay compared to the other two links that both have equal delay. Node 1 has 100 Mbps of demand that it can obtain from 2 as well as 3. In addition, there is 20 Mbps of demand at node 1 which it can obtain only from 2.  We assume that the aggregate demand at a node consists of a large number of user-initiated connections. When content can be downloaded from multiple locations, users initiate parallel TCP connections and the throughputs along paths in a parallel TCP connection are inversely proportional to the path delays. The TE scheme is assumed to be OSPF-based, i.e., shortest-path routing using configured link weights and traffic split equally among multiple paths with equal weights.

@@

%The following two observations motivate us to revisit ISP traffic engineering:
%\begin{itemize}
%	\item
%	The use of link-utilization based metrics may not be a good predictor of application performance that depends on other factors such as queuing delay, access and backbone link capacity, etc. 
%	\item
%	Traffic matrices form the basis of evaluating traffic engineering schemes, but application-level adaptation can change these traffic matrices itself in response to an updated routing.
%\end{itemize}

%The new direction we pursue is to shift focus from link-utxilzation based metrics to application-level metrics such as TCP file download times, and to account for the effect of application-level adaptation in evaluating traffic engineering schemes. In Chapter \ref{ch:beyondmlu}, we provide a comparison of traffic engineering schemes focusing on user-perceived metrics such as file download times, and VoIP call quality. In Chapter \ref{ch:beyondmlu} and Chapter \ref{ch:ncdn}, we evaluate TE schemes while accounting for the effect of application-level adaptation and show that this interaction helps simpler TE schemes, such as oblivious TE or OSPF TE, perform closer to the optimal TE strategy in terms of user-perceived metrics as well as TE metrics.

\subsection{Does content placement flexibility obviate sophisticated traffic engineering?}

\textbf{Example 2: Demand-aware content placement} To appreciate how placement can shape traffic, consider the simple example in Figure~\ref{fig:NetworkExample}. Node $C$ has an object in its cache that is requested by end-users at nodes $A$ and $D$. Suppose that one unit of traffic needs to be routed from $C$ to $A$ and $0.5$ units  from $C$ to $D$ to satisfy the demand for that object. The routing that achieves the minimum MLU of $0.5$ to serve the demanded object is shown in the figure. Note that the routing that achieves the MLU of 0.5 is not possible with a simple, \unplanned\ protocol like InverseCap as that would route all the traffic demand from $C$ to $A$ via $B$, resulting in an MLU of 1. Thus, a (\planned) traffic engineering scheme is necessary to achieve an MLU of 0.5.

\begin{figure}[h]
	\centering
	\includegraphics[width=2in]{ncdnpaper/ncdn-example}
	\caption{A simple \ncp\ example}
	\vspace{-.3in}
	\label{fig:NetworkExample}
\end{figure}

@@

In this example, content placement flexibility reduces network cost and enables simpler routing. But, what happens for real content and traffic traces? How effective are simple placement schemes towards optimizing user-perceived performance and network cost?

% As earlier examples show, content placement can be a more powerful degree of freedom than request redirection and network routing.

% What is the relative importance of placement vs routing? Can content placement obviate the need for sophisticated traffic engineering schemes?

\subsection{Do Network CDNs need joint optimization of content delivery and traffic engineering?}

\textbf{How well do simple demand oblivous compare against demand-aware?}

\textbf{An NCDN can potentially jointly optimize content delivery and traffic engineeing, but do such sophisticated schemes yield benefits from a CDN?}

A Network CDN (NCDN) is an ISP than owns and operates a CDN over its infrastructures to deliver content to its end users. As NCDNs manage both the underlying network and content delivey, the objectives and the techniques available to a network CDNs are not the same from a traditional ISP or a tradidional CDN. Our work is motivated by the following questions:

\begin{itemize}
	\item 
	Is a demand-aware or a demand-oblivious content delivery and traffic engineering scheme more effective for a network CDN?
	\item
	An NCDN can potentially jointly optimize content delivery and traffic engineeing, but do such sophisticated schemes yield benefits from a CDN?
\end{itemize}

To our knowledge, this thesis presents the first study of techniques for a managing a NCDN based on real network topology and content access traces. Our result show the effectiveness of simple demand-oblivious schemes for content delivery and traffic engineering in achieving performance close to an ideal scheme. Further, they show that a demand-aware joint optimization performs poorly due to frequnt churn in content workloads and change in access patterns.

\eat{
\section{Illustrating the interaction between content placement and traffic engineering}

@@

}

### 2015-07-10 tepaper/section1.tex
Background: updated motivation, and Chapter intro subsection planning for Beyond MLU & NCDN

\chapter{Traffic Engineering: An Application-Centric Comparison}
\label{ch:beyondmlu}
Traditionally, traffic engineering (TE) has been studied as an optimization problem that takes as input a traffic matrix (TM) and seeks to compute routes so as to minimize a network cost function. The cost function is intended to capture the severity of congestion hotpsots based on link utilization levels. For example, the most widely used cost function, MLU, is simply the utilization of the most utilized link in the network \cite{COPE,TEXCP,MultiTM,Cohen}; others  sum over all links a convex function of their utilization  (so as to penalize highly utilized links more)  \cite{fortz2000internet,fortz2002traffic}. There are two implicit assumptions underlying this line of work. First, maintaining low link utilization improves user-perceived application performance under typical load conditions. Second, maintaining low link utilization increases the effective capacity of the network by enabling it to accommodate unexpected surges in the traffic demand.

>>>>

\chapter{Traffic Engineering: An Application-Centric Comparison}
\label{ch:beyondmlu}

\section{Introduction}

\subsection{Shortcomings of link utilization metrics}

\subsection{Challenges in evaluating TE schemes}

\subsection{Results from an application-centric evaluation}

\subsection{Chapter organization}

Traditionally, traffic engineering (TE) has been studied as an optimization problem that takes as input a traffic matrix (TM) and seeks to compute routes so as to minimize a network cost function. The cost function is intended to capture the severity of congestion hotpsots based on link utilization levels. For example, the most widely used cost function, MLU, is simply the utilization of the most utilized link in the network \cite{COPE,TEXCP,MultiTM,Cohen}; others  sum over all links a convex function of their utilization  (so as to penalize highly utilized links more)  \cite{fortz2000internet,fortz2002traffic}. There are two implicit assumptions underlying this line of work. First, maintaining low link utilization improves user-perceived application performance under typical load conditions. Second, maintaining low link utilization increases the effective capacity of the network by enabling it to accommodate unexpected surges in the traffic demand.
